---
title: "Multiple MAF R2 (Affy)"
author: "Helena Davies"
date: "27/04/2022"
output: html_document
---
---
title: "Multiple MAF and R2"
author: "Helena Davies"
date: "20/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Delete everything in your global environment
```{r Delete everything in your global environment}
remove(list = ls())
```

Call in functions script
```{r Source functions}
source(file = "/Users/helenadavies/werk/ilovedata/scripts/functions/add_numeric.R")
source(file = "/Users/helenadavies/werk/ilovedata/scripts/functions/remove_duplicates.R")
source(file = "/Users/helenadavies/werk/ilovedata/scripts/functions/package_check.R")
```

Use package_check to install and load dependencies
Load tidyverse last
```{r Install load dependencies}
packages <- c("summarytools",
               "psych",
               "tidyverse")
package_check(packages)

```

Retrieve the recent date
```{r Recent date}
date = Sys.Date()
date
```

```{bash eval=FALSE, include=FALSE}
# Download file 
rsync k19047584@login.rosalind.kcl.ac.uk:/mnt/lustre/groups/gwased/data_raw/CH_GLAD_NBR_ED/imputed/pfile/CHAffy_GLADv2_NBRv1_EUR_imputed_biallele_rsid_SEX_exl_maf001_R2_3_chr1.pvar /Users/helenadavies/Desktop 
```

# Read in data
```{r read in data}
dat <- read.table("/Users/helenadavies/King's College London/MT-Translational Neuropsychiatric Genomics - Helena_Davies_PhD - Helena_Davies_PhD/PGC_Freeze3_cases/data_cleaned/genetics/CHAffy_GLADv2_NBRv1_EUR_imputed_biallele_rsid_SEX_exl_maf001_R2_3_chr1.pvar")

dat <- dat %>%
  select(chr_position = V3,
         AF_IMPUTED_MAF_R2 = V7
         )
```

```{r split column into multiple columns}
dat_split1 <- dat %>%
  separate(AF_IMPUTED_MAF_R2,
           into = c("AF_IMPUTED",
                    "MAF_R2"),
           sep = "MAF=",
           remove = TRUE
           )

dat_split2 <- dat_split1 %>%
  separate(MAF_R2,
           into = c("MAF",
                    "R2"),
           sep = ";R2=",
           remove = TRUE
           )

dat_split3 <- dat_split2 %>%
  separate(MAF,
           into = c("MAF1",
                    "MAF2",
                    "MAF3"),
           sep = ",",
           remove = TRUE
           )

dat_split4 <- dat_split3 %>%
  separate(R2,
           into = c("R2_1",
                    "R2_2",
                    "R2_3"),
           sep = ",",
           remove = TRUE
           )
```

```{r convert MAF and R2 to numeric}
dat_split4$MAF1 <- as.numeric(dat_split4$MAF1)
dat_split4$MAF2 <- as.numeric(dat_split4$MAF2)
dat_split4$MAF3 <- as.numeric(dat_split4$MAF3)
dat_split4$R2_1 <- as.numeric(dat_split4$R2_1)
dat_split4$R2_2 <- as.numeric(dat_split4$R2_2)
dat_split4$R2_3 <- as.numeric(dat_split4$R2_3)

# Select final columns
dat_split_final <- dat_split4 %>%
  select(chr_position,
         MAF1,
         MAF2,
         MAF3,
         R2_1,
         R2_2,
         R2_3)
```

# First, filter out variants with only one MAF and/or only R2 value
```{r create new variable of difference between MAF}
dat_split_final %>%
  nrow() # 1389416

dat_split_final <- dat_split_final %>%
  filter((!is.na(MAF2) &
         !is.na(MAF3)) |
           (!is.na(R2_2) |
            !is.na(R2_3))  )

dat_split_final %>%
  nrow() # 1044388

1389416 - 1044388 # = 345028 variants have only one MAF and/or one R2 value
```

# Second, make new variables of difference between MAF1 and MAF2, and R2_1 and R2_2
```{r new variable of difference}
# MAF 1 and 2 difference
dat_split_final <- dat_split_final %>%
  mutate(MAF_diff1_2 = 
           case_when(!is.na(MAF1) ~ MAF1 - MAF2))

# Check
dat_split_final %>%
  descr(MAF_diff1_2)

# MAF 1 and 3 difference
dat_split_final <- dat_split_final %>%
  mutate(MAF_diff1_3 = 
           case_when(!is.na(MAF1) ~ MAF1 - MAF3))

# Check
dat_split_final %>%
  descr(MAF_diff1_3)

# MAF 2 and 3 difference
dat_split_final <- dat_split_final %>%
  mutate(MAF_diff2_3 = 
           case_when(!is.na(MAF1) ~ MAF2 - MAF3))

# Check
dat_split_final %>%
  descr(MAF_diff2_3)


# R2 1 and 2 difference
dat_split_final <- dat_split_final %>%
  mutate(R2_diff1_2 = 
           case_when(!is.na(R2_1) ~ R2_1 - R2_2))

# Check
dat_split_final %>%
  descr(R2_diff1_2)

# R2 1 and 3 difference
dat_split_final <- dat_split_final %>%
  mutate(R2_diff1_3 = 
           case_when(!is.na(R2_1) ~ R2_1 - R2_3))

# Check
dat_split_final %>%
  descr(R2_diff1_3)


# R2 2 and 3 difference
dat_split_final <- dat_split_final %>%
  mutate(R2_diff2_3= 
           case_when(!is.na(R2_1) ~ R2_2 - R2_3))

# Check
dat_split_final %>%
  descr(R2_diff2_3)
```

# Next, make the differences absolute values
```{r make diffs absolute}
dat_split_final$MAF_diff1_2 <- abs(dat_split_final$MAF_diff1_2)
dat_split_final$R2_diff1_2 <- abs(dat_split_final$R2_diff1_2)
dat_split_final$MAF_diff1_3<- abs(dat_split_final$MAF_diff1_3)
dat_split_final$R2_diff1_3 <- abs(dat_split_final$R2_diff1_3)
dat_split_final$MAF_diff2_3 <- abs(dat_split_final$MAF_diff2_3)
dat_split_final$R2_diff2_3 <- abs(dat_split_final$R2_diff2_3)
```

# Finally, filter by MAFs and/or R2 differences of under 0.05
```{r filter out diffs over 0.05}
dat_split_final %>%
  nrow() # 1044388

dat_split_final <- dat_split_final %>%
  filter(MAF_diff1_2 <= 0.05 &
         MAF_diff1_3 <= 0.05 &
         MAF_diff2_3 <= 0.05 &
           
         R2_diff1_2 <= 0.05 &
         R2_diff1_3 <= 0.05 &
         R2_diff2_3 <= 0.05 )

dat_split_final %>%
  nrow() # 744106

1044388 - 744106 # = 300,282
# TOTAL VARIANTS LOST 
1389416 - 744106 # = 645,310 Variants lost in total
```
# Save the chromosome position in separate file
```{r save chr_position}
dat_split_final$chr_position <- as.factor(dat_split_final$chr_position)

dat_split_final %>%
  select(chr_position) 
  write.csv("/Users/helenadavies/King's College London/MT-Translational Neuropsychiatric Genomics - Helena_Davies_PhD - Helena_Davies_PhD/PGC_Freeze3_cases/data_cleaned/genetics/CH_Affy_chr1_keep_factor.csv", row.names = FALSE, quote = FALSE)
```











