---
title: "UKB file formatting"
author: "Helena Davies"
date: "26/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Delete everything in your global environment
```{r Delete everything in your global environment}
remove(list = ls())
```

# Load packages
```{r load packages}
#install.packages("devtools")
library(devtools)

#install.packages("DescTools")
library(DescTools)

#install.packages("stringi")
#library(stringi)

#install.packages("tidyverse")
library(tidyverse)
```

# Read in data
## Case
```{r read in case data}
case_dat <- readRDS(file = "/scratch/groups/ukbiobank/usr/helena_d/UKB_PGCED_Freeze3_final_files/UKB_PGCED3_only_case_data_070322.rds")

# Check
case_dat %>%
  colnames()

case_dat %>%
  nrow()
```

## Control
```{r read in control data}
control_dat <- readRDS(file = "/scratch/groups/ukbiobank/usr/helena_d/UKB_PGCED_Freeze3_final_files/UKB_PGCED3_control_data_100322.rds")

# Check
control_dat %>%
  colnames()

control_dat %>%
  nrow()

# Create control variable
control_dat$control <- 1

# Check
control_dat %>%
  count(control)

# Select variables
control_dat <- control_dat %>%
  select(eid,
         control)
```

# Merge all data
```{r merge all data}
dat <- dplyr::full_join(control_dat,
                         case_dat,
                         by ="eid")

# Check
dat %>%
  colnames()
```

# Controls = 0, case = 1
Regenie: for binary traits, must be coded as 0=control, 1=case, NA=missing unless using --1
```{r coding AN for regenie}
dat <- dat %>%
  mutate(AN =
           case_when(AN_case == 1 ~ 1, # Case
                     
                     control == 1 ~ 0, # Controls 
                     
                     TRUE ~ NA_real_ # Neither controls nor cases (i.e., BE cases)
                     )
  )

# Check
dat %>%
  count(AN)
```

```{r coding BE_broad for regenie}
dat <- dat %>%
  mutate(BE_broad =
           case_when(BE_broad == 1 ~ 1, # Case
                     
                      control == 1 ~ 0, # Controls 
                     
                     TRUE ~ NA_real_ # Neither controls nor cases (i.e., AN cases)
                     )
  )

# Check
dat %>%
  count(BE_broad)
```

```{r coding BE_broad for regenie}
dat <- dat %>%
  mutate(BE_narrow =
           case_when(BE_narrow == 1 ~ 1, # Case
                     
                      control == 1 ~ 0, # Controls 
                     
                     TRUE ~ NA_real_ # Neither controls nor cases (i.e., AN cases or BE broad)
                     )
  )

# Check
dat %>%
  count(BE_narrow)
```

```{r coding FID for regenie}
dat <- dat %>%
  mutate(FID =
           case_when(!is.na(eid)  ~ eid
                     )
  )

# Check
length(unique(dat$FID))
```

```{r select variables}
dat <- dat %>%
  select(FID,
         IID = eid,
         AN,
         BE_broad,
         BE_narrow)

# Final checks
head(dat)

dat %>%
  count(AN)

dat %>%
  count(BE_broad)

dat %>%
  count(BE_narrow)
```

```{r save as txt file space deliminated}
write.table(dat,
            file = "/scratch/groups/ukbiobank/usr/helena_d/UKB_PGCED3_final_case_control_dat_100322.txt",
            col.names=T,
            row.names = F,
            quote=F,
            sep=" ")
```

# Find centre information
```{r read in data for centre info}
#Point to the project directory
project_dir <- "/scratch/datasets/ukbiobank/ukb18177"

#Read the project field-to-name “field finder” file
f <- ukbkings::bio_field(project_dir)

# inspect the variable metadata
head(f)
glimpse(f)

#display the number of baskets included.
f %>%
distinct(basket)
```

# Search for variables required 
```{r search for centre information}
f %>%
select(name, field) %>%
page(method = "print")
```

# Add their field codes to a file, one per line, no header. You can page through the file.
```{r}
strings <- c("eid",
             "54") # UKB assessment centre

f %>%
select(field, name) %>%
filter(str_detect(field,
                  paste(strings, collapse = "|"))) %>%
  bio_field_add("small_field_subset.txt")
```

# Read required fields and save as an rds file in your user directory. Argument out should be a path to your UKB user directory
```{r}
bio_phen(
 project_dir,
 field = "small_field_subset.txt",
 out = "small_phenotype_subset"
)
```

# Check the size of your file and read in your dataset
```{r}
system("ls -lh small_phenotype_subset.rds")
df <- readRDS("small_phenotype_subset.rds")
```

Rename columns from the default UKB field names to the descriptive names used in the field-to-name “field finder” name column.

0	Initial assessment visit (2006-2010) at which participants were recruited and consent given
1	First repeat assessment visit (2012-13)
2	Imaging visit (2014+)
3	First repeat imaging visit (2019+)


11012	Barts
11021	Birmingham
11011	Bristol
11008	Bury
11003	Cardiff
11024	Cheadle (revisit)
11020	Croydon
11005	Edinburgh
11004	Glasgow
11018	Hounslow
11010	Leeds
11016	Liverpool
11001	Manchester
11017	Middlesborough
11009	Newcastle
11013	Nottingham
11002	Oxford
11007	Reading
11014	Sheffield
10003	Stockport (pilot)
11006	Stoke
11022	Swansea
11023	Wrexham
11025	Cheadle (imaging)
11026	Reading (imaging)
11027	Newcastle (imaging)
11028	Bristol (imaging)
```{r rename columns using bio_rename}
# Initial assessment
df <- df %>%
  mutate(centre_initial_assessment =
           case_when(`54-0.0_ukb37667` == "11012" ~ "Barts",
                     `54-0.0_ukb37667` == "11012" ~  "Barts",
                     `54-0.0_ukb37667` == "11021" ~  "Birmingham",
                     `54-0.0_ukb37667` == "11011" ~  "Bristol",
                     `54-0.0_ukb37667` == "11008" ~  "Bury",
                     `54-0.0_ukb37667` == "11003" ~  "Cardiff",
                     `54-0.0_ukb37667` == "11024" ~  "Cheadle (revisit)",
                     `54-0.0_ukb37667` == "11020" ~  "Croydon",
                     `54-0.0_ukb37667` == "11005" ~  "Edinburgh",
                     `54-0.0_ukb37667` == "11004" ~  "Glasgow",
                     `54-0.0_ukb37667` == "11018" ~  "Hounslow",
                     `54-0.0_ukb37667` == "11010" ~  "Leeds",
                     `54-0.0_ukb37667` == "11016" ~  "Liverpool",
                     `54-0.0_ukb37667` == "11001" ~  "Manchester",
                     `54-0.0_ukb37667` == "11017" ~  "Middlesborough",
                     `54-0.0_ukb37667` == "11009" ~  "Newcastle",
                     `54-0.0_ukb37667` == "11013" ~  "Nottingham",
                     `54-0.0_ukb37667` == "11002" ~  "Oxford",
                     `54-0.0_ukb37667` == "11007" ~  "Reading",
                     `54-0.0_ukb37667` == "11014" ~  "Sheffield",
                     `54-0.0_ukb37667` == "10003" ~  "Stockport (pilot)",
                     `54-0.0_ukb37667` == "11006" ~  "Stoke",
                     `54-0.0_ukb37667` == "11022" ~  "Swansea",
                     `54-0.0_ukb37667` == "11023" ~  "Wrexham",
                     `54-0.0_ukb37667` == "11025" ~  "Cheadle (imaging)",
                     `54-0.0_ukb37667` == "11026" ~  "Reading (imaging)",
                     `54-0.0_ukb37667` == "11027" ~  "Newcastle (imaging)",
                     `54-0.0_ukb37667` == "11028" ~  "Bristol (imaging)"
           )
  )

# Check
df %>%
  count(centre_initial_assessment)

# First repeat assessment                    
df <- df %>%
  mutate(centre_first_repeat_assessment =
           case_when(`54-1.0_ukb37667` == "11012" ~ "Barts",
                     `54-1.0_ukb37667` == "11012" ~  "Barts",
                     `54-1.0_ukb37667` == "11021" ~  "Birmingham",
                     `54-1.0_ukb37667` == "11011" ~  "Bristol",
                     `54-1.0_ukb37667` == "11008" ~  "Bury",
                     `54-1.0_ukb37667` == "11003" ~  "Cardiff",
                     `54-1.0_ukb37667` == "11024" ~  "Cheadle (revisit)",
                     `54-1.0_ukb37667` == "11020" ~  "Croydon",
                     `54-1.0_ukb37667` == "11005" ~  "Edinburgh",
                     `54-1.0_ukb37667` == "11004" ~  "Glasgow",
                     `54-1.0_ukb37667` == "11018" ~  "Hounslow",
                     `54-1.0_ukb37667` == "11010" ~  "Leeds",
                     `54-1.0_ukb37667` == "11016" ~  "Liverpool",
                     `54-1.0_ukb37667` == "11001" ~  "Manchester",
                     `54-1.0_ukb37667` == "11017" ~  "Middlesborough",
                     `54-1.0_ukb37667` == "11009" ~  "Newcastle",
                     `54-1.0_ukb37667` == "11013" ~  "Nottingham",
                     `54-1.0_ukb37667` == "11002" ~  "Oxford",
                     `54-1.0_ukb37667` == "11007" ~  "Reading",
                     `54-1.0_ukb37667` == "11014" ~  "Sheffield",
                     `54-1.0_ukb37667` == "10003" ~  "Stockport (pilot)",
                     `54-1.0_ukb37667` == "11006" ~  "Stoke",
                     `54-1.0_ukb37667` == "11022" ~  "Swansea",
                     `54-1.0_ukb37667` == "11023" ~  "Wrexham",
                     `54-1.0_ukb37667` == "11025" ~  "Cheadle (imaging)",
                     `54-1.0_ukb37667` == "11026" ~  "Reading (imaging)",
                     `54-1.0_ukb37667` == "11027" ~  "Newcastle (imaging)",
                     `54-1.0_ukb37667` == "11028" ~  "Bristol (imaging)"
           )
  )
             
# Check
df %>%
  count(centre_first_repeat_assessment)           


# First repeat assessment                    
df <- df %>%
  mutate(centre_imaging_assessment =
           case_when(`54-2.0_ukb37667` == "11012" ~ "Barts",
                     `54-2.0_ukb37667` == "11012" ~  "Barts",
                     `54-2.0_ukb37667` == "11021" ~  "Birmingham",
                     `54-2.0_ukb37667` == "11011" ~  "Bristol",
                     `54-2.0_ukb37667` == "11008" ~  "Bury",
                     `54-2.0_ukb37667` == "11003" ~  "Cardiff",
                     `54-2.0_ukb37667` == "11024" ~  "Cheadle (revisit)",
                     `54-2.0_ukb37667` == "11020" ~  "Croydon",
                     `54-2.0_ukb37667` == "11005" ~  "Edinburgh",
                     `54-2.0_ukb37667` == "11004" ~  "Glasgow",
                     `54-2.0_ukb37667` == "11018" ~  "Hounslow",
                     `54-2.0_ukb37667` == "11010" ~  "Leeds",
                     `54-2.0_ukb37667` == "11016" ~  "Liverpool",
                     `54-2.0_ukb37667` == "11001" ~  "Manchester",
                     `54-2.0_ukb37667` == "11017" ~  "Middlesborough",
                     `54-2.0_ukb37667` == "11009" ~  "Newcastle",
                     `54-2.0_ukb37667` == "11013" ~  "Nottingham",
                     `54-2.0_ukb37667` == "11002" ~  "Oxford",
                     `54-2.0_ukb37667` == "11007" ~  "Reading",
                     `54-2.0_ukb37667` == "11014" ~  "Sheffield",
                     `54-2.0_ukb37667` == "10003" ~  "Stockport (pilot)",
                     `54-2.0_ukb37667` == "11006" ~  "Stoke",
                     `54-2.0_ukb37667` == "11022" ~  "Swansea",
                     `54-2.0_ukb37667` == "11023" ~  "Wrexham",
                     `54-2.0_ukb37667` == "11025" ~  "Cheadle (imaging)",
                     `54-2.0_ukb37667` == "11026" ~  "Reading (imaging)",
                     `54-2.0_ukb37667` == "11027" ~  "Newcastle (imaging)",
                     `54-2.0_ukb37667` == "11028" ~  "Bristol (imaging)"
           )
  )
             
# Check
df %>%
  count(centre_imaging_assessment)       

df <- df %>%
  mutate(centre_first_repeat_imaging_assessment =
           case_when(`54-3.0_ukb37667` == "11012" ~ "Barts",
                     `54-3.0_ukb37667` == "11012" ~  "Barts",
                     `54-3.0_ukb37667` == "11021" ~  "Birmingham",
                     `54-3.0_ukb37667` == "11011" ~  "Bristol",
                     `54-3.0_ukb37667` == "11008" ~  "Bury",
                     `54-3.0_ukb37667` == "11003" ~  "Cardiff",
                     `54-3.0_ukb37667` == "11024" ~  "Cheadle (revisit)",
                     `54-3.0_ukb37667` == "11020" ~  "Croydon",
                     `54-3.0_ukb37667` == "11005" ~  "Edinburgh",
                     `54-3.0_ukb37667` == "11004" ~  "Glasgow",
                     `54-3.0_ukb37667` == "11018" ~  "Hounslow",
                     `54-3.0_ukb37667` == "11010" ~  "Leeds",
                     `54-3.0_ukb37667` == "11016" ~  "Liverpool",
                     `54-3.0_ukb37667` == "11001" ~  "Manchester",
                     `54-3.0_ukb37667` == "11017" ~  "Middlesborough",
                     `54-3.0_ukb37667` == "11009" ~  "Newcastle",
                     `54-3.0_ukb37667` == "11013" ~  "Nottingham",
                     `54-3.0_ukb37667` == "11002" ~  "Oxford",
                     `54-3.0_ukb37667` == "11007" ~  "Reading",
                     `54-3.0_ukb37667` == "11014" ~  "Sheffield",
                     `54-3.0_ukb37667` == "10003" ~  "Stockport (pilot)",
                     `54-3.0_ukb37667` == "11006" ~  "Stoke",
                     `54-3.0_ukb37667` == "11022" ~  "Swansea",
                     `54-3.0_ukb37667` == "11023" ~  "Wrexham",
                     `54-3.0_ukb37667` == "11025" ~  "Cheadle (imaging)",
                     `54-3.0_ukb37667` == "11026" ~  "Reading (imaging)",
                     `54-3.0_ukb37667` == "11027" ~  "Newcastle (imaging)",
                     `54-3.0_ukb37667` == "11028" ~  "Bristol (imaging)"
           )
  )
             
# Check
df %>%
  count(centre_first_repeat_imaging_assessment)       
```

```{r save centre data}
df %>%
  select(ID = eid,
        centre_initial_assessment,
        centre_first_repeat_assessment,
        centre_imaging_assessment,
         centre_first_repeat_imaging_assessment
         ) %>%
  write.table(file = "/scratch/groups/ukbiobank/usr/helena_d/UKB_PGCED3_centre_dat_100322.txt",
            row.names = F,
            col.names=T,
            quote=F,
            sep=" ")
```

# Now, covariate file formatting
## Read in data
```{r read in control data}
covariates_dat <- readRDS(file = "/scratch/groups/ukbiobank/usr/helena_d/UKB_PGCED3_covariates.rds")

# Check
covariates_dat %>%
  colnames()

covariates_dat %>%
  nrow()
```

```{r covariate file formatting}
covariates_dat <- covariates_dat %>%
  mutate(FID =
           case_when(!is.na(eid)  ~ eid
                     )
  )

# Check
length(unique(covariates_dat$FID))

colnames(covariates_dat)
```

```{r select variables}
covariates_dat <- covariates_dat %>%
  select(FID,
         IID = eid,
         sex,
         age_at_recruitment,
         bmi_at_initial_recruitment)

# Final checks
head(covariates_dat)
```

```{r save as txt file space deliminated}
write.table(covariates_dat,
            file = "/scratch/groups/ukbiobank/usr/helena_d/UKB_PGCED3_final_covariates_dat.txt",
            row.names = F,
            col.names=T,
            quote=F,
            sep=" ")
```