---
title: "UKB PGCED controls"
author: "Helena Davies"
date: "28/02/2022"
output: html_document
---

# Set up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Delete everything in your global environment
```{r Delete everything in your global environment}
remove(list = ls())
```

# Load packages
```{r load packages}
#install.packages("devtools")
library(devtools)

# Install Ken's UKB KCL package
#devtools::install_github("kenhanscombe/ukbkings",
#          dependencies = TRUE,
#           force = TRUE)
#
library(ukbkings)

#install.packages("DescTools")
library(DescTools)

#install.packages("stringi")
#library(stringi)

#install.packages("tidyverse")
library(tidyverse)

#install.packages("tidyr")
library(tidyr)
```

# Controls
# You need a file with required fields, one per line, no header.
```{r}
#Point to the project directory
project_dir <- "/scratch/datasets/ukbiobank/ukb18177"

#Read the project field-to-name “field finder” file
f <- ukbkings::bio_field(project_dir)

# inspect the variable metadata
head(f)
glimpse(f)

#display the number of baskets included.
f %>%
distinct(basket)
```

# Search for variables required 
```{r}
f %>%
select(name, field) %>%
page(method = "print")
```

# Add their field codes to a file, one per line, no header. You can page through the file.
```{r}
strings <- c("eid",
             "20544", # for MHQ - psychological overeating or binge eating
             "20002") # for non-cancer illness code - anorexia/bulimia/other eating disorder

f %>%
select(field, name) %>%
filter(str_detect(field,
                  paste(strings, collapse = "|"))) %>%
  bio_field_add("small_field_subset.txt")
```

# Read required fields and save as an rds file in your user directory. Argument out should be a path to your UKB user directory
```{r}
bio_phen(
 project_dir,
 field = "small_field_subset.txt",
 out = "small_phenotype_subset"
)
```

# Check the size of your file and read in your dataset
```{r}
system("ls -lh small_phenotype_subset.rds")
df <- readRDS("small_phenotype_subset.rds")
```

Rename columns from the default UKB field names to the descriptive names used in the field-to-name “field finder” name column.
```{r rename columns using bio_rename}
df <- bio_rename(df, f)
nrow(df)

colnames(df)
```

# first, read in case data
```{r read in case data}
case_dat <- readRDS(file = "/scratch/groups/ukbiobank/usr/helena_d/UKB_PGCED_Freeze3_final_files/UKB_PGCED3_only_case_data_070322.rds")

case_dat <- case_dat %>% 
  filter(BE_narrow == 1 |
           BE_broad == 1 |
           AN_case == 1) # 1958
```

# Filter by ID (i.e., select controls based on not being a case)
```{r filter control data by case data}
`%!in%` = Negate(`%in%`)

control_dat <- df %>%
  filter(eid %!in% case_dat$eid)

# Check 
control_dat %>% # 500581 (502539 [total UKB] - 1958 [PGCED cases])
  nrow() 
``` 

```{r remove not needed data to save memory}
rm(case_dat)
gc()
```

# Need to apply additional filters
i.e., Also filter by ANY other eating disorder [include filtering by cases] or binge eating


## Death data (need to also filter out for 'other eating disorder' death cases)
Death by AN is included in cases (and therefore automatically not included in controls.) No other death by eating disorders are endorsed other than 'F50.8 Other eating disorders' n = 1

- Death register - Health-related outcomes
- Contributory (secondary) causes of death: ICD10
    - F50.8 Other eating disorders n = 1
```{r DEATH read in data}
death_cause_diskframe <- bio_record(project_dir, record = "death_cause")

death_cause <- death_cause_diskframe %>%
    select(eid,
           cause_icd10
           ) %>%
    collect()

death_cause %>%
  nrow()
```

```{r death register not cases not controls}
# Death due to any other ED
death_data <- death_cause %>%
  mutate(death_due_to_any_other_ED =
           case_when(cause_icd10 == "F508" ~ 1
            ))

death_data %>%
  count(death_due_to_any_other_ED)  # 1
```
  
```{r filter out cases}
death_data_not_controls <- death_data %>%
  filter(death_due_to_any_other_ED == 1)

death_data_not_controls %>%
  count(death_due_to_any_other_ED)

control_dat <- control_dat %>%
  filter(eid %!in% death_data_not_controls$eid)

# Check
control_dat %>%
  nrow() # 500580 (500581 [UKB control data] - 1 [death due to ED])
```

```{r remove not needed data to save memory}
rm(death_data_not_controls)
rm(death_data)
rm(death_cause)
gc()
```

## Non-cancer illness code, self-reported
Medical conditions - Verbal interview - Assessment Centre
neurology/eye/psychiatry --> psychological/psychiatric problem --> substance abuse/dependency --> anorexia/bulimia/other eating disorder (n = 410)
https://biobank.ndph.ox.ac.uk/showcase/coding.cgi?id=6
```{r filter out non-cancer illness code verbal interview not cases not controls}
# Initial assessment
df <- df %>%
  mutate(anorexia_bulimia_other_eating_disorder0_1 = 
           case_when(
           `20002-0.0_ukb50079` == "1470" |
           `20002-0.1_ukb50079` == "1470" |
           `20002-0.2_ukb50079` == "1470" |
           `20002-0.3_ukb50079` == "1470" |
           `20002-0.4_ukb50079` == "1470" |
           `20002-0.5_ukb50079` == "1470" |
           `20002-0.6_ukb50079` == "1470" |
           `20002-0.7_ukb50079` == "1470" |
           `20002-0.8_ukb50079` == "1470" |
           `20002-0.9_ukb50079` == "1470" |
           `20002-0.10_ukb50079`  == "1470" |
           `20002-0.11_ukb50079`  == "1470" |
           `20002-0.12_ukb50079`  == "1470" |
           `20002-0.13_ukb50079`  == "1470" |
           `20002-0.14_ukb50079`  == "1470" |
           `20002-0.15_ukb50079`  == "1470" |
           `20002-0.16_ukb50079`  == "1470" |
           `20002-0.17_ukb50079`  == "1470" |
           `20002-0.18_ukb50079`  == "1470" |
           `20002-0.19_ukb50079`  == "1470" |
           `20002-0.20_ukb50079`  == "1470" |
           `20002-0.21_ukb50079`  == "1470" |
           `20002-0.22_ukb50079`  == "1470" |
           `20002-0.22_ukb50079`  == "1470" |
           `20002-0.22_ukb50079`  == "1470" |
           `20002-0.23_ukb50079`  == "1470" |
           `20002-0.24_ukb50079`  == "1470" |
           `20002-0.25_ukb50079`  == "1470" |
           `20002-0.26_ukb50079`  == "1470" |
           `20002-0.27_ukb50079`  == "1470" |
           `20002-0.28_ukb50079`  == "1470" |
           `20002-0.29_ukb50079`  == "1470" |
           `20002-0.30_ukb50079`  == "1470" |
           `20002-0.31_ukb50079`  == "1470" |
           `20002-0.32_ukb50079` == "1470" |
           `20002-0.33_ukb50079` == "1470" |
           `20002-1.0_ukb50079` == "1470" |
           `20002-1.1_ukb50079` == "1470" |
           `20002-1.2_ukb50079` == "1470" |
           `20002-1.3_ukb50079` == "1470" |
           `20002-1.4_ukb50079` == "1470" |
           `20002-1.5_ukb50079` == "1470" |
           `20002-1.6_ukb50079` == "1470" |
           `20002-1.7_ukb50079` == "1470" |
           `20002-1.8_ukb50079` == "1470" |
           `20002-1.9_ukb50079` == "1470" |
           `20002-1.10_ukb50079`  == "1470" |
           `20002-1.11_ukb50079`  == "1470" |
           `20002-1.12_ukb50079`  == "1470" |
           `20002-1.13_ukb50079`  == "1470" |
           `20002-1.14_ukb50079`  == "1470" |
           `20002-1.15_ukb50079`  == "1470" |
           `20002-1.16_ukb50079`  == "1470" |
           `20002-1.17_ukb50079`  == "1470" |
           `20002-1.18_ukb50079`  == "1470" |
           `20002-1.19_ukb50079`  == "1470" |
           `20002-1.20_ukb50079`  == "1470" |
           `20002-1.21_ukb50079`  == "1470" |
           `20002-1.22_ukb50079`  == "1470" |
           `20002-1.22_ukb50079`  == "1470" |
           `20002-1.22_ukb50079`  == "1470" |
           `20002-1.23_ukb50079`  == "1470" |
           `20002-1.24_ukb50079`  == "1470" |
           `20002-1.25_ukb50079`  == "1470" |
           `20002-1.26_ukb50079`  == "1470" |
           `20002-1.27_ukb50079`  == "1470" |
           `20002-1.28_ukb50079`  == "1470" |
           `20002-1.29_ukb50079`  == "1470" |
           `20002-1.30_ukb50079`  == "1470" |
           `20002-1.31_ukb50079`  == "1470" |
           `20002-1.32_ukb50079` == "1470" |
           `20002-1.33_ukb50079` == "1470" ~ 1))
          
df %>%
  count(anorexia_bulimia_other_eating_disorder0_1) # 379
            
# Second assessment              
df <- df %>%
  mutate(anorexia_bulimia_other_eating_disorder2= 
           case_when(             
           `20002-2.0_ukb50079` == "1470" |
           `20002-2.1_ukb50079` == "1470" |
           `20002-2.2_ukb50079` == "1470" |
           `20002-2.3_ukb50079` == "1470" |
           `20002-2.4_ukb50079` == "1470" |
           `20002-2.5_ukb50079` == "1470" |
           `20002-2.6_ukb50079` == "1470" |
           `20002-2.7_ukb50079` == "1470" |
           `20002-2.8_ukb50079` == "1470" |
           `20002-2.9_ukb50079` == "1470" |
           `20002-2.10_ukb50079`  == "1470" |
           `20002-2.11_ukb50079`  == "1470" |
           `20002-2.12_ukb50079`  == "1470" |
           `20002-2.13_ukb50079`  == "1470" |
           `20002-2.14_ukb50079`  == "1470" |
           `20002-2.15_ukb50079`  == "1470" |
           `20002-2.16_ukb50079`  == "1470" |
           `20002-2.17_ukb50079`  == "1470" |
           `20002-2.18_ukb50079`  == "1470" |
           `20002-2.19_ukb50079`  == "1470" |
           `20002-2.20_ukb50079`  == "1470" |
           `20002-2.21_ukb50079`  == "1470" |
           `20002-2.22_ukb50079`  == "1470" |
           `20002-2.22_ukb50079`  == "1470" |
           `20002-2.22_ukb50079`  == "1470" |
           `20002-2.23_ukb50079`  == "1470" |
           `20002-2.24_ukb50079`  == "1470" |
           `20002-2.25_ukb50079`  == "1470" |
           `20002-2.26_ukb50079`  == "1470" |
           `20002-2.27_ukb50079`  == "1470" |
           `20002-2.28_ukb50079`  == "1470" |
           `20002-2.29_ukb50079`  == "1470" |
           `20002-2.30_ukb50079`  == "1470" |
           `20002-2.31_ukb50079`  == "1470" |
           `20002-2.32_ukb50079` == "1470" |
           `20002-2.33_ukb50079` == "1470" ~ 1))
          
 df %>%
  count(anorexia_bulimia_other_eating_disorder2) # 20
               
 # Third assessment     
 df <- df %>%
  mutate(anorexia_bulimia_other_eating_disorder3= 
           case_when(             
            `20002-3.0_ukb50079` == "1470" |
           `20002-3.1_ukb50079` == "1470" |
           `20002-3.2_ukb50079` == "1470" |
           `20002-3.3_ukb50079` == "1470" |
           `20002-3.4_ukb50079` == "1470" |
           `20002-3.5_ukb50079` == "1470" |
           `20002-3.6_ukb50079` == "1470" |
           `20002-3.7_ukb50079` == "1470" |
           `20002-3.8_ukb50079` == "1470" |
           `20002-3.9_ukb50079` == "1470" |
           `20002-3.10_ukb50079`  == "1470" |
           `20002-3.11_ukb50079`  == "1470" |
           `20002-3.12_ukb50079`  == "1470" |
           `20002-3.13_ukb50079`  == "1470" |
           `20002-3.14_ukb50079`  == "1470" |
           `20002-3.15_ukb50079`  == "1470" |
           `20002-3.16_ukb50079`  == "1470" |
           `20002-3.17_ukb50079`  == "1470" |
           `20002-3.18_ukb50079`  == "1470" |
           `20002-3.19_ukb50079`  == "1470" |
           `20002-3.20_ukb50079`  == "1470" |
           `20002-3.21_ukb50079`  == "1470" |
           `20002-3.22_ukb50079`  == "1470" |
           `20002-3.22_ukb50079`  == "1470" |
           `20002-3.22_ukb50079`  == "1470" |
           `20002-3.23_ukb50079`  == "1470" |
           `20002-3.24_ukb50079`  == "1470" |
           `20002-3.25_ukb50079`  == "1470" |
           `20002-3.26_ukb50079`  == "1470" |
           `20002-3.27_ukb50079`  == "1470" |
           `20002-3.28_ukb50079`  == "1470" |
           `20002-3.29_ukb50079`  == "1470" |
           `20002-3.30_ukb50079`  == "1470" |
           `20002-3.31_ukb50079`  == "1470" |
           `20002-3.32_ukb50079` == "1470" |
           `20002-3.33_ukb50079` == "1470" ~ 1))
   
df %>%
  count(anorexia_bulimia_other_eating_disorder3) # 5
         
# Overall
df <- df %>%
  mutate(anorexia_bulimia_other_eating_disorder= 
           case_when(  
             anorexia_bulimia_other_eating_disorder0_1 == 1 | 
              anorexia_bulimia_other_eating_disorder2 ==  1 |
             anorexia_bulimia_other_eating_disorder3 == 1 ~ 1))  
 
    
df %>%
  count(anorexia_bulimia_other_eating_disorder) # 397
```

```{r filter out cases}
noncancer_illness_not_controls <- df %>%
  filter(anorexia_bulimia_other_eating_disorder == 1)

noncancer_illness_not_controls %>%
  count(anorexia_bulimia_other_eating_disorder)


unique(length(noncancer_illness_not_controls$eid))

control_dat <- control_dat %>%
  filter(eid %!in% noncancer_illness_not_controls$eid)

# Check
control_dat %>%
  nrow() # 500406 (500580 [UKB control data] - 174 [noncancer illness code of 'anorexia/bulimia/other eating disorder' that had not already been excluded])
```

```{r remove not needed data to save memory}
rm(noncancer_illness_not_controls)
gc()
```

## MHQ data
### Psychological over-eating or binge-eating
```{r MHQ psych overeating or binge eating}
df <- df %>%
  mutate(psych_overeating_or_BE_MHQ =
           case_when(
          mental_health_problems_ever_diagnosed_by_a_professional_f20544_0_1 == "13" |
           mental_health_problems_ever_diagnosed_by_a_professional_f20544_0_2 == "13" |
           mental_health_problems_ever_diagnosed_by_a_professional_f20544_0_3 == "13" |
           mental_health_problems_ever_diagnosed_by_a_professional_f20544_0_4 == "13" |
           mental_health_problems_ever_diagnosed_by_a_professional_f20544_0_5 == "13" |
           mental_health_problems_ever_diagnosed_by_a_professional_f20544_0_6 == "13" |
           mental_health_problems_ever_diagnosed_by_a_professional_f20544_0_7 == "13" |
           mental_health_problems_ever_diagnosed_by_a_professional_f20544_0_8 == "13" |
           mental_health_problems_ever_diagnosed_by_a_professional_f20544_0_9 == "13" |
           mental_health_problems_ever_diagnosed_by_a_professional_f20544_0_10 == "13" |
           mental_health_problems_ever_diagnosed_by_a_professional_f20544_0_11 == "13" |
           mental_health_problems_ever_diagnosed_by_a_professional_f20544_0_12 == "13" |
           mental_health_problems_ever_diagnosed_by_a_professional_f20544_0_13 == "13" |
           mental_health_problems_ever_diagnosed_by_a_professional_f20544_0_14 == "13" |
           mental_health_problems_ever_diagnosed_by_a_professional_f20544_0_15 == "13" |
           mental_health_problems_ever_diagnosed_by_a_professional_f20544_0_16 == "13" ~ 1
     )
  )


df %>%
  count(psych_overeating_or_BE_MHQ) # 707
```

Filter out cases 
```{r MHQ filter out cases}
MHQ_dat_not_controls <- df %>%
  filter(psych_overeating_or_BE_MHQ == 1)

control_dat <- control_dat %>%
  filter(eid %!in% MHQ_dat_not_controls$eid)

# Check
control_dat %>%
  nrow() # 499822 (500406 [UKB control data] - 584 [MHQ code of 'psychological overeating/binge eating' that had not already been excluded])
```

```{r remove not needed data to save memory}
rm(MHQ_dat_not_controls)
gc()
```

## HES data
```{r HES read in diagnosis data}
hesin_diag_diskframe <- bio_record(project_dir, record = "hesin_diag")

hesin_diag <- hesin_diag_diskframe %>%
    select(eid,
           diag_icd9,
           diag_icd10,
           ) %>%
    collect()

nrow(hesin_diag)
colnames(hesin_diag)
```

### Diagnoses - ICD9
Decided NOT to exclude ICD9 code of 'anorexia' (7830) as very likely this is cancer-related anorexia, or (7836) polyphagia because this is not a full eating disorder or binge eating 

### Diagnoses - ICD10
Decided NOT to exclude ICD10 code of 'Vomiting associated with other psychological disturbances' (F505) ICD10 code of 'Abnormal weight loss' (R634) or 'Abnormal weight gain' (R635) or 'Insufficient intake of food and water due to self neglect' (R636) or  Polyphagia (R632) or Feeding difficulties and mismanagement (R633) as this does not necessarily meet diagnostic threshold for an eating disorder
```{r ICD10 diagnoses cases not controls}
# F50.8 Other eating disorders
hesin_diag <- hesin_diag %>%
  mutate(ICD10_other_EDs =
           case_when(diag_icd10 == "F508" ~ 1
            ))

hesin_diag %>%
  count(ICD10_other_EDs) 

# F50.9 Eating disorder, unspecified
hesin_diag <- hesin_diag %>%
  mutate(ICD10_eating_disorder_unspecified =
           case_when(diag_icd10 == "F509" ~ 1
            ))

hesin_diag %>%
  count(ICD10_eating_disorder_unspecified) 
```

Filter out cases 
```{r HES filter out cases}
HES_dat_not_controls <- hesin_diag %>%
  filter(ICD10_other_EDs == 1 |
         ICD10_eating_disorder_unspecified == 1 )

control_dat <- control_dat %>%
  filter(eid %!in% HES_dat_not_controls$eid)

# Check
control_dat %>%
  nrow() # 499777  (499822 [UKB control data] - 45 [HES code of an eating disorder that had not already been excluded])
```

```{r remove not needed data to save memory}
rm(HES_dat_not_controls)
rm(hesin_diag)
gc()
```

## GP data
Controls: Any individual who does not have a history of binge eating (narrow or broad definition) AND does not have a history of any eating disorder.

NB: All GP codes for anorexia nervosa and binge eating counted as cases have already been excluded from controls.
```{r read in GP data controls}
gp_clinical_diskframe <- bio_record(project_dir, record = "gp_clinical")

gp_clinical <- gp_clinical_diskframe %>%
    select(eid,
           read_2,
           read_3
           
           ) %>%
    collect()

head(gp_clinical)
```

### GP read 2 code: ‘History of any eating disorder’
```{r GP read code 2 History of any eating disorder}
# 1612. = Appetite loss - anorexia; Anorexia symptom; Loss of appetite - symptom
gp_clinical <- gp_clinical %>%
  mutate(read2_appetite_loss_anorexia_anorexia_symptom_loss_of_appetite =
           case_when(read_2 == "1612." ~ 1
           )
  )

gp_clinical %>%
  count(read2_appetite_loss_anorexia_anorexia_symptom_loss_of_appetite)  # 203


# 8HTN. = Referral to eating disorders clinic
gp_clinical <- gp_clinical %>%
  mutate(read2_referral_to_ED_clinic =
           case_when(read_2 == "8HTN." ~ 1
           )
  )

gp_clinical %>%
  count(read2_referral_to_ED_clinic) # 12

# 9Nk9. = Seen in eating disorder clinic
gp_clinical <- gp_clinical %>%
  mutate(read2_seen_in_ED_clinic =
           case_when(read_2 == "9Nk9." ~ 1
           )
  )

gp_clinical %>%
  count(read2_seen_in_ED_clinic) # 4

# E275. = Other and unspecified non-organic eating disorders
gp_clinical <- gp_clinical %>%
  mutate(read2_other_unspec_nonorganic_EDs =
           case_when(read_2 == "E275." ~ 1
           )
  )

gp_clinical %>%
  count(read2_other_unspec_nonorganic_EDs) # 12

# E2750 = Unspecified non-organic eating disorder
gp_clinical <- gp_clinical %>%
  mutate(read2_unspec_nonorganic_EDs =
           case_when(read_2 == "E2750" ~ 1
           )
  )

gp_clinical %>%
  count(read2_unspec_nonorganic_EDs) # 3

# E2752 = Pica
gp_clinical <- gp_clinical %>%
  mutate(read2_pica =
           case_when(read_2 == "E2752" ~ 1
           )
  )

gp_clinical %>%
  count(read2_pica) # 2

# E275y = Other specified non-organic eating disorder
gp_clinical <- gp_clinical %>%
  mutate(read2_other_unspec_nonorganic_EDs_2 =
           case_when(read_2 == "E275y" ~ 1
           )
  )

gp_clinical %>%
  count(read2_other_unspec_nonorganic_EDs_2) # 0


# E275z = Non-organic eating disorder NOS
gp_clinical <- gp_clinical %>%
  mutate(read2_nonorganic_EDs_NOS =
           case_when(read_2 == "E275z" ~ 1
           )
  )

gp_clinical %>%
  count(read2_nonorganic_EDs_NOS) # 2


# Eu50. = [X]Eating disorders
gp_clinical <- gp_clinical %>%
  mutate(read2_XEating_disorders =
           case_when(read_2 == "Eu50." ~ 1
           )
  )

gp_clinical %>%
  count(read2_XEating_disorders) # 64


# Eu501 = [X]Atypical anorexia nervosa
gp_clinical <- gp_clinical %>%
  mutate(read2_Xatypical_anorexia_nervosa =
           case_when(read_2 == "Eu501" ~ 1
           )
  )

gp_clinical %>%
  count(read2_Xatypical_anorexia_nervosa) # 2


# Eu50y = [X]Other eating disorders; [X]Pica in adults; [X]Psychogenic loss of appetite
gp_clinical <- gp_clinical %>%
  mutate(read2_XOther_eating_disorders_pica_psychogenic_loss_appetite =
           case_when(read_2 == "Eu50y" ~ 1
           )
  )

gp_clinical %>%
  count(read2_XOther_eating_disorders_pica_psychogenic_loss_appetite) # 1

# Eu50z = [X]Eating disorder, unspecified
gp_clinical <- gp_clinical %>%
  mutate(read2_XEating_disorder_unspec =
           case_when(read_2 == "Eu50z" ~ 1
           )
  )

gp_clinical %>%
  count(read2_XEating_disorder_unspec) # 4


# Fy05. = Nocturnal sleep-related eating disorder
gp_clinical <- gp_clinical %>%
  mutate(read2_nocturnal_sleeprelated_eating_disorder =
           case_when(read_2 == "Fy05." ~ 1
           )
  )

gp_clinical %>%
  count(read2_nocturnal_sleeprelated_eating_disorder) # 0


# R030. = [D]Anorexia (even if excl criteria not met)
gp_clinical <- gp_clinical %>%
  mutate(read2_DAnorexia =
           case_when(read_2 == "R030." ~ 1
           )
  )

gp_clinical %>%
  count(read2_DAnorexia) # 54


# R030z = [D]Anorexia NOS (even if excl criteria not met)
gp_clinical <- gp_clinical %>%
  mutate(read2_DAnorexia_NOS =
           case_when(read_2 == "R030z" ~ 1
           )
  )

gp_clinical %>%
  count(read2_DAnorexia_NOS) # 0
```

### GP read 3 code: ‘History of any eating disorder’
```{r GP read code 3 history of any eating disorder}
# 1612. = Appetite loss: [anorexia] or [anorexia symptom]; Anorexia symptom; Appetite loss - anorexia; Loss of appetite; Loss of appetite - symptom
gp_clinical <- gp_clinical %>%
  mutate(read3_appetite_loss_anorexia =
           case_when(read_3 == "1612." ~ 1
           )
  )

gp_clinical %>%
  count(read3_appetite_loss_anorexia) 

# R0360 = [D]Excessive eating; [D]Bulimia NOS
gp_clinical <- gp_clinical %>%
  mutate(read3_excessive_eating_bulimia_NOS =
           case_when(read_3 == "R0360" ~ 1
           )
  )

gp_clinical %>%
  count(read3_excessive_eating_bulimia_NOS) 

# E275. = Other and unspecified non-organic eating disorders
gp_clinical <- gp_clinical %>%
  mutate(read3_other_unspec_nonorganic_eating_disorders =
           case_when(read_3 == "E275." ~ 1
           )
  )

gp_clinical %>%
  count(read3_other_unspec_nonorganic_eating_disorders) 


# E2750 = Unspecified non-organic eating disorder
gp_clinical <- gp_clinical %>%
  mutate(read3_unspecified_nonorganic_eating_disorder =
           case_when(read_3 == "E2750" ~ 1
           )
  )

gp_clinical %>%
  count(read3_unspecified_nonorganic_eating_disorder) 


# E2752 = Pica; eats abnormal objects
gp_clinical <- gp_clinical %>%
  mutate(read3_pica_or_eats_abnormal_objects =
           case_when(read_3 == "E2752" ~ 1
           )
  )

gp_clinical %>%
  count(read3_pica_or_eats_abnormal_objects) 

# Eu50. = [X]Eating disorders
gp_clinical <- gp_clinical %>%
  mutate(read3_XEating_disorders =
           case_when(read_3 == "Eu50." ~ 1
           )
  )

gp_clinical %>%
  count(read3_XEating_disorders)


# Eu50z = [X]Eating disorder, unspecified
gp_clinical <- gp_clinical %>%
  mutate(read3_XEating_disorders_unspec =
           case_when(read_3 == "Eu50z" ~ 1
           )
  )

gp_clinical %>%
  count(read3_XEating_disorders_unspec)


# Xa2hW = Dietary advice for eating disorder
gp_clinical <- gp_clinical %>%
  mutate(read3_dietary_advice_for_ED =
           case_when(read_3 == "Xa2hW" ~ 1
           )
  )

gp_clinical %>%
  count(read3_dietary_advice_for_ED)

# XaEC2 = Eating disorder counselling
gp_clinical <- gp_clinical %>%
  mutate(read3_eating_disorder_counselling =
           case_when(read_3 == "XaEC2" ~ 1
           )
  )

gp_clinical %>%
  count(read3_eating_disorder_counselling)


# XaPBE = Seen in eating disorder clinic
gp_clinical <- gp_clinical %>%
  mutate(read3_seen_in_eating_disorder_clinic =
           case_when(read_3 == "XaPBE" ~ 1
           )
  )

gp_clinical %>%
  count(read3_seen_in_eating_disorder_clinic)



# R030. = [D]Anorexia 
gp_clinical <- gp_clinical %>%
  mutate(read3_DAnorexia =
           case_when(read_3 == "R030." ~ 1
           )
  )

gp_clinical %>%
  count(read3_DAnorexia)


# R030z = [D]Anorexia NOS 
gp_clinical <- gp_clinical %>%
  mutate(read3_DAnorexia_NOS =
           case_when(read_3 == "R030z" ~ 1
           )
  )

gp_clinical %>%
  count(read3_DAnorexia_NOS)


# X00Sx = Eating disorder
gp_clinical <- gp_clinical %>%
  mutate(read3_eating_disorder =
           case_when(read_3 == "X00Sx" ~ 1
           )
  )

gp_clinical %>%
  count(read3_eating_disorder)

# X00Sz = Atypical anorexia nervosa
gp_clinical <- gp_clinical %>%
  mutate(read3_atypical_anorexia_nervosa =
           case_when(read_3 == "X00Sz" ~ 1
           )
  )

gp_clinical %>%
  count(read3_atypical_anorexia_nervosa)


# XM07X = Lack of appetite; Anorexia; Off food; No appetite; Anorexic (even if excl criteria not met)
gp_clinical <- gp_clinical %>%
  mutate(read3_lack_of_appetite_anorexia_off_food_no_appetite =
           case_when(read_3 == "XM07X" ~ 1
           )
  )

gp_clinical %>%
  count(read3_lack_of_appetite_anorexia_off_food_no_appetite)


# .1612 = Anorexia symptom; Appetite loss: [anorexia] or [anorexia symptom]; Appetite loss; anorexia; Loss of appetite; Loss of appetite - symptom
gp_clinical <- gp_clinical %>%
  mutate(read3_appetite_loss_anorexia_2 =
           case_when(read_3 == ".1612" ~ 1
           )
  )

gp_clinical %>%
  count(read3_appetite_loss_anorexia_2)


# Eu501 = Atypical anorexia nervosa
gp_clinical <- gp_clinical %>%
  mutate(read3_atypical_anorexia_nervosa_2 =
           case_when(read_3 == "Eu501" ~ 1
           )
  )

gp_clinical %>%
  count(read3_atypical_anorexia_nervosa_2)


# XE1bQ = (Eating disorder: [bulimia - nonorganic] or [psychogenic overeating] or [NOS - psychogenic]) or (pica) or (infant feeding problem); Pica; Infant feeding problem; Bulimia - nonorganic; Psychogen. overeating; Eating disorder NOS - psychog.; Feeding problem-infant
gp_clinical <- gp_clinical %>%
  mutate(read3_nonorganic_bulimia_psychogenic_overeating_pica_infant_feeding_problem =
           case_when(read_3 == "XE1bQ" ~ 1
           )
  )

gp_clinical %>%
  count(read3_nonorganic_bulimia_psychogenic_overeating_pica_infant_feeding_problem)


# XafjT = Eating disorder co-occurrent with diabetes mellitus type 1; Diabulimia; ED-DMT1 - Eating disorder-diabetes mellitus type 1
gp_clinical <- gp_clinical %>%
  mutate(read3_eating_disorder_cooccurrent_with_diabetes =
           case_when(read_3 == "XafjT" ~ 1
           )
  )

gp_clinical %>%
  count(read3_eating_disorder_cooccurrent_with_diabetes)


# .E4D4 = (Eating disorder: [bulimia - nonorganic] or [psychogenic overeating] or [NOS - psychogenic]) or (pica) or (infant feeding problem); Pica; Infant feeding problem; Bulimia - nonorganic; Psychogen. overeating; Eating disorder NOS - psychog; Feeding problem-infant
gp_clinical <- gp_clinical %>%
  mutate(read3_nonorganic_bulimia_psychogenic_overeating_pica_infant_feeding_problem_2 =
           case_when(read_3 == ".E4D4" ~ 1
           )
  )

gp_clinical %>%
  count(read3_nonorganic_bulimia_psychogenic_overeating_pica_infant_feeding_problem_2)


# E275y = Other specified non-organic eating disorder
gp_clinical <- gp_clinical %>%
  mutate(read3_other_unspec_nonorganic_eating_disorder =
           case_when(read_3 == "E275y" ~ 1
           )
  )

gp_clinical %>%
  count(read3_other_unspec_nonorganic_eating_disorder) # 1


# E275z = Non-organic eating disorder NOS
gp_clinical <- gp_clinical %>%
  mutate(read3_nonorganic_eating_disorder_NOS =
           case_when(read_3 == "E275z" ~ 1
           )
  )

gp_clinical %>%
  count(read3_nonorganic_eating_disorder_NOS)


# Eu50y = [X] Eating disorders: [other] or [pica in adults] or [psychogenic loss of appetite]; [X]Other eating disorders; [X] Pica in adults; [X] Psychogenic loss of appetite; [X]Pica in adults; [X]Psychogenic loss of appetite
gp_clinical <- gp_clinical %>%
  mutate(read3_XEating_disorder_pica_psychogenic_loss_appetite =
           case_when(read_3 == "Eu50y" ~ 1
           )
  )

gp_clinical %>%
  count(read3_XEating_disorder_pica_psychogenic_loss_appetite) # 0

# .R340 = [D]Excessive eating; [D]Bulimia NOS
gp_clinical <- gp_clinical %>%
  mutate(read3_DExcessive_eating_bulimia_NOS =
           case_when(read_3 == ".R340" ~ 1
           )
  )

gp_clinical %>%
  count(read3_DExcessive_eating_bulimia_NOS) # 0
```

### GP read 2 code: Binge eating
```{r GP read code 2 binge eating}
# 1FN.. = Emotional eating
gp_clinical <- gp_clinical %>%
  mutate(read2_emotional_eating =
           case_when(read_2 == "1FN.." ~ 1
           )
  )

gp_clinical %>%
  count(read2_emotional_eating) # 1
```

### GP read code 3 binge eating
```{r GP read code 3 binge eating}
# X767Z = Binge eating; Bingeing; Bouts of overeating; Episodes of overeating; Binges
gp_clinical <- gp_clinical %>%
  mutate(read3_binge_eating_bingeing_bouts_overeating_binges =
           case_when(read_3 == "X767Z" ~ 1
           )
  )

gp_clinical %>%
  count(read3_binge_eating_bingeing_bouts_overeating_binges) # 33

# .1FF. = Binge eating; Bingeing; Bouts of overeating; Episodes of overeating’ Binges
gp_clinical <- gp_clinical %>%
  mutate(read3_binge_eating_bingeing_bouts_overeating_binges_2 =
           case_when(read_3 == ".1FF." ~ 1
           )
  )

gp_clinical %>%
  count(read3_binge_eating_bingeing_bouts_overeating_binges_2) # 0

# 1FF.. = Binge eating; Bingeing; Bouts of overeating; Episodes of overeating’ Binges
gp_clinical <- gp_clinical %>%
  mutate(read3_binge_eating_bingeing_bouts_overeating_binges_3 =
           case_when(read_3 == "1FF.." ~ 1
           )
  )

gp_clinical %>%
  count(read3_binge_eating_bingeing_bouts_overeating_binges_3) # 0

# Xab0U = Emotional eating
gp_clinical <- gp_clinical %>%
  mutate(read3_emotional_eating =
           case_when(read_3 == "Xab0U" ~ 1
           )
  )

gp_clinical %>%
  count(read3_emotional_eating) # 0

# 1FN.. = Emotional eating
gp_clinical <- gp_clinical %>%
  mutate(read3_emotional_eating_2 =
           case_when(read_3 == "1FN.." ~ 1
           )
  )

gp_clinical %>%
  count(read3_emotional_eating_2) # 0
```

### Filter out cases
```{r filter out cases}
gp_not_controls <- gp_clinical %>%
  filter(
    # Read 2: Any eating disorder
    read2_appetite_loss_anorexia_anorexia_symptom_loss_of_appetite == 1 |
    read2_referral_to_ED_clinic == 1 |
    read2_seen_in_ED_clinic == 1 |
    read2_other_unspec_nonorganic_EDs == 1 |
    read2_unspec_nonorganic_EDs == 1 |
    read2_pica == 1 |
    read2_other_unspec_nonorganic_EDs_2 == 1 |
    read2_nonorganic_EDs_NOS == 1 |
    read2_XEating_disorders == 1 |
    read2_Xatypical_anorexia_nervosa == 1 |
    read2_XOther_eating_disorders_pica_psychogenic_loss_appetite == 1 |
    read2_XEating_disorder_unspec == 1 |
    read2_nocturnal_sleeprelated_eating_disorder == 1 |
    read2_DAnorexia == 1 |
    read2_DAnorexia_NOS == 1 |
     
       # Read 3: Any eating disorder
      read3_appetite_loss_anorexia == 1 |
      read3_excessive_eating_bulimia_NOS == 1 |
      read3_other_unspec_nonorganic_eating_disorders == 1 |
      read3_unspecified_nonorganic_eating_disorder == 1 |
      read3_pica_or_eats_abnormal_objects == 1 |
      read3_XEating_disorders == 1 |
      read3_XEating_disorders_unspec == 1 |
      read3_dietary_advice_for_ED == 1 |
      read3_eating_disorder_counselling == 1 |
      read3_seen_in_eating_disorder_clinic == 1 |
      read3_DAnorexia == 1 |
      read3_DAnorexia_NOS == 1 |
      read3_eating_disorder == 1 |
      read3_atypical_anorexia_nervosa == 1 |
      read3_lack_of_appetite_anorexia_off_food_no_appetite == 1 |
      read3_appetite_loss_anorexia_2 == 1 |
      read3_atypical_anorexia_nervosa_2 == 1 |
      read3_nonorganic_bulimia_psychogenic_overeating_pica_infant_feeding_problem == 1 |
      read3_eating_disorder_cooccurrent_with_diabetes == 1 |
      read3_nonorganic_bulimia_psychogenic_overeating_pica_infant_feeding_problem_2 == 1 |
      read3_other_unspec_nonorganic_eating_disorder == 1 |
      read3_nonorganic_eating_disorder_NOS == 1 |
      read3_XEating_disorder_pica_psychogenic_loss_appetite == 1 |
      read3_DExcessive_eating_bulimia_NOS == 1 |
     
       # Read 2: Binge eating
      read2_emotional_eating == 1 |
      
      # Read 3: Binge eating
      read3_binge_eating_bingeing_bouts_overeating_binges == 1 |
      read3_binge_eating_bingeing_bouts_overeating_binges_2 == 1 |
      read3_binge_eating_bingeing_bouts_overeating_binges_3 == 1 |
      read3_emotional_eating == 1 |
      read3_emotional_eating_2 == 1 
    ) 


control_dat <- control_dat %>%
  filter(eid %!in% gp_not_controls$eid) 

control_dat %>%
  nrow() # 499308 (499777 - 469 [participants with a GP code of an eating disorder or binge who have not already been excluded])
```

# Save
```{r save control data}
saveRDS(object = control_dat, file = paste0("/scratch/groups/ukbiobank/usr/helena_d/UKB_PGCED_Freeze3_final_files/UKB_PGCED3_control_data_100322.rds"))
```



